Machine learning high multiplicity matrix elements for electron-positron and hadron-hadron colliders

### 2022-12-29 qcd.tex
for submission pt2

\section{Divergent structures}\label{sec:divergences}
    The computation of matrix elements essentially reduces
    to evaluating Feynman diagrams which are analytical expressions
    built up from the Feynman rules of a theory (c.f. QCD Feynman rules
    Figure \ref{fig:qcd_feynman_rules}). In evaluating certain
    topologies of Feynman diagrams, integral expressions
    containing unconstrained momenta will give rise to
    singularities. The divergences
    associated with high energy modes are called ultraviolet
    (UV) divergences, and on the opposite end of the energy
    spectrum, there are infrared (IR) divergences associated
    with low energy modes.

    The de facto methods to alleviate these
    divergences are through renormalisation
    for UV divergences, and through subtraction
    schemes for IR divergences. Both of these
    methods will be discussed in this section.

@@

    In Section~\ref{sec:antenna_functions} we will give
    a brief overview of antenna functions where these one-loop
    factorisation formulae were used to obtain universal singular functions
    at the one-loop level for squared matrix elements. These
    antenna functions are then applied in the context of NLO QCD
    K-factor emulation for electron-positron annihilation in Chapter~\ref{chapter:fame2}.

@@

        A key feature of these kernels are that they smoothly interpolate
        between the soft and collinear limits, meaning they do not double count
        any limit. Additionally, they only contain divergences in the
        $p_{i}p_{j} \rightarrow 0$ limit and not for any other pair of
        momenta. In the soft and collinear limits, the dipole function
        $\mathcal{D}_{ij,k}$ correctly reproduces the matrix element factorisation
        behaviour seen in Eqs. (\ref{eqn:soft_factorisation}) and (\ref{eqn:F_collinear_factorisation}),
        respectively. In particular, $\boldsymbol{V}_{ij,k}$ becomes
        proportional to the eikonal factor and Altarelli-Parisi splitting
        functions in these respective limits.

@@

        The universality of infrared divergent structure
        extends to massive quarks \cite{Catani:2000ef}.
        In Reference \cite{Catani:2002hc} the authors derive dipole
        functions for massive partons which
        are used in Chapter~\ref{chapter:fame3}, but will
        not be written here as the procedure of constructing
        the massive dipole functions is similar to their massless counterparts.

@@

    For the one-loop antenna functions, there are three
    different functions corresponding to the leading colour,
    sub-leading colour, and closed quark loop contributions.
    It is sufficient to consider the leading colour
    structure in this thesis as the remaining contributions
    are sub-leading. Ignoring sub-leading terms in
    colour is equivalent to taking the large $N_{c}$
    limit. For QCD where $N_{c} = 3$, neglecting
    sub-leading colour terms amounts to an
    approximately 10\% effect. (cite here)

>>>>

\section{Divergent structures}\label{sec:divergences}
    The computation of matrix elements essentially reduces
    to evaluating Feynman diagrams which are analytical expressions
    built up from the Feynman rules of a theory (c.f. QCD Feynman rules
    in Figure \ref{fig:qcd_feynman_rules}). In evaluating certain
    topologies of Feynman diagrams, integral expressions
    containing unconstrained momenta will give rise to
    singularities. The divergences
    associated with high energy modes are called ultraviolet
    (UV) divergences, and on the opposite end of the energy
    spectrum, there are infrared (IR) divergences associated
    with low energy modes.

    The de facto methods to alleviate these
    divergences are through renormalisation
    for UV divergences, and through subtraction
    schemes for IR divergences. Both of these
    methods will be discussed in this chapter.

@@

    In Section~\ref{sec:antenna_functions} we will give
    a brief overview of antenna functions where these one-loop
    factorisation formulae were used to obtain universal singular functions
    at the one-loop level for squared matrix elements. These
    antenna functions are then applied in the context of NLO QCD
    k-factor emulation for electron-positron annihilation in Chapter~\ref{chapter:fame2}.

@@

        A key feature of these kernels is that they smoothly interpolate
        between the soft and collinear limits, meaning they do not double count
        any limit. Additionally, they only contain divergences in the
        $p_{i}p_{j} \rightarrow 0$ limit and not for any other pair of
        momenta. In the soft and collinear limits, the dipole function
        $\mathcal{D}_{ij,k}$ correctly reproduces the matrix element factorisation
        behaviour seen in Eqs. (\ref{eqn:soft_factorisation}) and (\ref{eqn:F_collinear_factorisation}),
        respectively. In particular, $\boldsymbol{V}_{ij,k}$ becomes
        proportional to the eikonal factor and Altarelli-Parisi splitting
        functions in these respective limits.

@@

        The universality of infrared divergent structure
        extends to massive quarks \cite{Catani:2000ef}.
        In Ref. \cite{Catani:2002hc} the authors derive dipole
        functions for massive partons which
        are used in Chapter~\ref{chapter:fame3}, but will
        not be written here as the procedure of constructing
        the massive dipole functions is similar to their massless counterparts.

@@

    For the one-loop antenna functions, there are three
    different functions corresponding to the leading colour,
    sub-leading colour, and closed quark loop contributions.
    It is sufficient to consider the leading colour
    structure in this thesis as the remaining contributions
    are sub-leading. Ignoring sub-leading terms in
    colour is equivalent to taking the large $N_{c}$
    limit. For QCD where $N_{c} = 3$, neglecting
    sub-leading colour terms amounts to an
    approximately 10\% effect.
